CudeleFS: Programmable Consistency and Durability\\in a Global Namespace

### 2017-03-31 paper/sections/0_abstract.tex
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

HPC developers debate abandoning POSIX because the synchronization and
serialization overheads of providing strong consistency and durability are too
costly -- and often uneccessary -- for their applications.  Unfortunately,
designing near-POSIX file systems excludes applications that rely on strong
consistency or durability, forcing developers to re-write their applications or
deploy them on a different system.  We present a file system and API that
allows clients to specify their consistency/durability requirements and assign
them to subtrees in the namespace, allowing administrators to optimize subtrees
within the same namespace for different workloads.  We draw conclusions about
the performance impact of unexplored consistency/durability metadata designs
and show that strong consistency can cause a 104\(\times\) slow down while merging
updates (7\(\times\) slow down) and maintaining durability (\(10\times\) slow
down) have more reasonable costs.

>>>>

HPC developers are abandoning POSIX because the synchronization and
serialization overheads of providing strong consistency and durability are too
costly -- and often uneccessary -- for their applications.  Unfortunately,
designing near-POSIX file systems excludes applications that rely on strong
consistency or durability, forcing developers to re-write their applications or
deploy them on a different system.  We present a file system and API that
lets clients to specify their consistency/durability requirements and assign
them to subtrees in the namespace, allowing administrators to optimize subtrees
within the same namespace for different workloads.  We draw conclusions about
the performance impact of unexplored consistency/durability metadata designs
and show that strong consistency can cause a 104\(\times\) slow down while merging
updates (7\(\times\) slow down) and maintaining durability (\(10\times\) slow
down) have more reasonable costs.

### 2017-03-31 paper/sections/1_introduction.tex
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is the problem?
File system metadata services in HPC have scalability problems. It has been
shown that HPC workloads, especially checkpoint-restart jobs, are resource
intensive primarily due to contention for the same directories and inodes ({\it
e.g.}, path traversal). Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is unreasonable. This problem is exacerbated
by current trends in HPC, where architectures are transitioning from complex
storage stacks with burst buffer, file system, object store, and tape tiers to
more simplified stackis with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
instead of data movement.

% What is HPC doing?
HPC's unique requirements for file systems ({\it e.g.}, fast creates) and well
defined workloads ({\it e.g.}, workflows) have given rise to file systems with
weaker consistency constraints than POSIX and less durability than programmers
are accustomed to.  Without these constraints, HPC systems can do more
client side processing with an emphasis on serverless metadata services.

One popular approach for relaxing consistency and durability is to ``decouple
the namespace", where clients lock the subtree they want exclusive access to so
other clients cannot interfere. This delayed merge ({\it i.e.} a form of
eventual consistency) and relaxed durability improves performance and
scalability by avoiding the costs of RPCs, synchronization, false sharing, and
serialization.  While the performance benefits are obvious for these users,
applications that rely on stronger consistency or durability guarantees must be
re-written or deployed on a different system. The consistency and durability
semantics for these systems is shown in Table~\ref{table:namespaces}.

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and durability policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

\begin{table}
\begin{tabular}{ r | l | l }
              & Decoupled & Global    \\
              & Namespace & Namespace \\\hline
  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
  Consistency & eventual & strong     \\
  Durability  & node local & global  \\
\end{tabular}
\caption{State-of-the-art systems in HPC improve file system metadata
performance by relaxing consistency and durability
guarantees.\label{table:namespaces}}
\end{table}

% What did we do
We propose subtree policies, an interface that lets future programmers control
how the storage system manages different parts of the file system namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

>>>>

% What is the problem?
File system metadata services in HPC have scalability problems. It has been
shown that HPC workloads are metadata resource intensive because administrative
tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs}, on large data sets leads to contention
for the same directories and inodes ({\it
e.g.}, path traversal). Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is unreasonable. This problem is exacerbated
by current trends in HPC, where architectures are transitioning from complex
storage stacks with burst buffer, file system, object store, and tape tiers to
more simplified stacks with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
because more requests end up hitting the same layer and latencies cannot be 
hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and durability policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, batch style jobs often do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client side processing and merge updates when the job is done. HPC
developers are turning to this non-POSIX solution because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because they wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
patterns).

% One example
One popular approach for relaxing consistency and durability is to ``decouple
the namespace", where clients lock the subtree they want exclusive access to as
a way to tell the file system that the subtree is important or may cause
resource contention in the near-future~\cite{grider:pdsw2015-marfs,
zheng:pdsw2015-deltafs, zheng:pdsw2014-batchfs, ren:sc2014-indexfs,
bent:slides-twotiers}. Then the file system can change its internal structure
to optimize performance. For example, the file system could enter a mode that
prevents other clients from interferring with the decoupled directory.  This
delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
durability improves performance and scalability by avoiding the costs of RPCs,
synchronization, false sharing, and serialization.  The consistency and
durability semantics for these systems is shown in the table in
Figure~\ref{fig:subtree-policies} . While the performance benefits are obvious
for these users, applications that rely on the file system's guarantees must be
deployed on an entirely different system or re-written to coordinate strong
consistency/durability themselves.

%\begin{table}
%\begin{tabular}{ r | l }
%  Subtree         & Example \\\hline
%  (1)   & \{Index, Batch\}FS~\cite{ren:sc2014-indexfs, zheng:pdsw2014-batchfs} \\
%  (2)   & \{Index, Ceph\}FS~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} \\
%  (3)   & RAMDisk \\
%  (4)   & DeltaFS~\cite{zheng:pdsw2015-deltafs} \\
%\end{tabular}
%
%\caption{State-of-the-art systems in HPC improve file system metadata
%performance by relaxing consistency and durability guarantees. Note that
%IndexFS also supports weak consistency with bulk inserts.
%\label{table:namespaces}} \end{table}

% What did we do
We propose subtree policies, an interface that lets future programmers control
the consistency and durability for subtrees in the file system namespace. For performance, one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, weak, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

### 2017-03-31 paper/sections/4_methodology.tex
paper: prune implementation section and make it readable
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete ({\it i.e.} the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``eventual" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); ; and updates in
``global" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
durability. These are just a few of the HPC examples.

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing operations like \texttt{lookup()}. The size of the inode cache is
configurable so as not to saturate the memory on the metadata server -- inodes
in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

>>>>

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete ({\it i.e.} the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); ; and updates in
``global" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``weak" consistency and ``local"
durability. These are just a few of the HPC examples.

@@

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing operations like \texttt{lookup()}.  For example, if a client or
metadata server is not caching the directory inode, all creates within that
directory will result in a lookup and a create request. If the directory inode
is cached then only the create needs to be sent.  The size of the inode cache
is configurable so as not to saturate the memory on the metadata server --
inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

### 2017-03-31 paper/sections/6_results.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: prune implementation section and make it readable
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsection{Eventual vs. Strong Consistency}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
\caption{The RPC per metadata update of ``Strong Consistency" has a large
overhead compared to the decoupled namespace strategy of ``Eventual
Consistency".\label{fig:slowdown-strong-eventual}}
\end{figure}

Figure~\ref{fig:slowdown-strong-eventual} shows the runtimes of eventual and
strong consistency based systems implemented on Cudele, normalized to the
runtime of the create mechanism (again, just creating files in the client's
in-memory journal). We scaled the number of files up to 100K which is the
maximum size of a directory by default in CephFS. We use the following
compositions from the mechanisms in Table~\ref{table:spectrum}:

@@

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: eventual vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: eventual, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

@@

\subsection{Isolation from Global Writes, Cost of Overwrites}
\subsection{Partial Directory Listings}

>>>>

\subsection{Eventual vs. Strong Consistency}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
\caption{The RPC per metadata update of ``Strong Consistency" has a large
overhead compared to the decoupled namespace strategy of ``Eventual
Consistency".\label{fig:slowdown-strong-weak}}
\end{figure}

Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of weak and
strong consistency based systems implemented on Cudele, normalized to the
runtime of the create mechanism (again, just creating files in the client's
in-memory journal). We scaled the number of files up to 100K which is the
maximum size of a directory by default in CephFS. We use the following
compositions from the mechanisms in Table~\ref{table:spectrum}:

@@

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

@@

\subsection{Isolation from Global Writes, Cost of Overwrites}
\subsection{Cost of Nonvolatile Apply}
% Why so slow?
Creating many files in the same directory would touch the same object but the
existing implementation results in this object being repeatedly pushed/pulled.

\subsection{Partial Directory Listings}
